How fluid-mechanical erosion creates anisotropic porous media

### 2022-06-29 docs/PhysicsPaper/StokesErosion.tex
NM

As a second way to quantify medium anisotropy, we measure the
tortuosity, $T_1$ and $T_2$, in the longitudinal and transverse
directions respectively. Figure~\ref{fig4}(a), shows how both vary over
the course of the simulation. As the bodies erode, $T_2$ grows
significantly larger than $T_1$, indicating that passive tracers must
follow more tortuous paths when traveling transversely and that
longitudinal paths are more direct. The ratio $(T_2-1)/(T_1-1)$ provides
a second measure of anisotropy to complement the permeability-dependent
quantity. We have chosen to subtract one in parenthesis so that both
numerator and denominator vanish when no bodies are present, as
analogous to resistivity. As seen in Figure~\ref{fig4}(b), the ratio $(T_2-1)/(T_1-1)$ grows significantly during erosion, up to a peak of nearly 8. This ratio can be decomposed into configurational and shape components, exactly as is done for the anisotropy (i.e.~by replacing a collection of eroded bodies with circles of the same areas and centers of mass). Figure~\ref{fig4}(b) shows that the shape component is greater for the majority of the simulation, but near the end of the simulation, the configurational component is competitive.

@@

Figure~\ref{fig5} shows the permeability and anisotropy measurements for the complete data set. In particular, for a given number of initial bodies ($M$ = 20, 40, 60, 80, or 100), the mean and standard deviation of each quantity are shown, with different values of $M$ shown by different colors. As seen in Figures~\ref{fig5}(a)--(b), the trends of increasing permeabilities are well grouped by the number of initial bodies. That is, permeability monotonically decreases with the number of bodies and the error bars shows little overlap between different values of $M$ until very late in the simulation when the bodies have nearly vanished.

>>>>

As a second way to quantify medium anisotropy, we measure the
tortuosity, $T_1$ and $T_2$, in the longitudinal and transverse
directions respectively. Figure~\ref{fig4}(a), shows how both vary over
the course of the simulation. As the bodies erode, $T_2$ grows
significantly larger than $T_1$, indicating that passive tracers must
follow more tortuous paths when traveling transversely compared to longitudinally. The ratio $(T_2-1)/(T_1-1)$ provides
a second measure of anisotropy to complement the permeability-dependent
quantity. We have chosen to subtract one in parenthesis so that both
numerator and denominator both vanish when no bodies are present, as
analogous to resistivity. As seen in Figure~\ref{fig4}(b), the ratio $(T_2-1)/(T_1-1)$ grows significantly during erosion, up to a peak of nearly 8. This ratio can be decomposed into configurational and shape components, exactly as is done for the anisotropy (i.e.~by replacing a collection of eroded bodies with circles of the same areas and centers of mass). Figure~\ref{fig4}(b) shows that the shape component is greater for the majority of the simulation, but near the end of the simulation, the configurational component is competitive.

@@

Figure~\ref{fig5} shows the permeability and anisotropy measurements for the complete data set. In particular, for a given number of initial bodies ($M$ = 20, 40, 60, 80, or 100), the mean of each quantity (permeability or anisotropy) is shown with the vertical error bars representing the standard deviation. Different initial numbers of bodies, $M$, are shown by different colors. As seen in Figures~\ref{fig5}(a)--(b), the trends of increasing permeabilities are well grouped by the number of initial bodies. That is, permeability monotonically decreases with the number of bodies and the error bars shows little overlap between different values of $M$ until very late in the simulation when the bodies have nearly vanished.
